Data exploration, quality control and statistical analysis of ChIP-exo experiments

### 2016-05-02 rmarkdown/paper.Rnw
abstract corrected

    ChIP-exo is a modification of the ChIP-Seq protocol for high
    resolution mapping of transcription factor binding sites. Although
    many aspect of the ChIP-exo data analysis are similar to those of
    ChIP-Seq, ChIP-exo presents a number of unique challenges. We
    present a quality control pipeline that analyzes a ChIP-exo and
    ChIP-nexus experiment's strand imbalance, enrichment and library
    complexity. Assessment of these characteristics are facilitated
    through diagnostic plots and summary statistics calculated over
    regions of the genome with varying levels of coverage.

    We systematically evaluated diverse aspects of ChIP-exo and found
    the following characteristics: First, ChIP-exo's background is
    quite different from ChIP-Seq's. Second, although often assumed in
    ChIP-exo data analysis methods, the ``peak pair'' assumptions does
    not hold locally in actual ChIP-exo data. Third, we for the first
    time compared Paired End (PE) ChIP-Seq with ChIP-exo and found
    that both protocols are comparable in resolutions and sensitivity
    for closely located binding events, but as the distance between
    binding events increases ChIP-exo shows higher sensitivity that PE
    ChIP-Seq. Finally, at fixed sequencing depths, ChIP-exo provides
    higher sensitivity, specificity and spatial resolution than PE
    ChIP-Seq.

  \end{abstract}

@@

>>>>

    ChIP-exo is a modification of the ChIP-Seq protocol for high
    resolution mapping of transcription factor binding sites. Although
    many aspects of the ChIP-exo data analysis are similar to those of
    ChIP-Seq, ChIP-exo presents a number of unique challenges. We
    present a quality control pipeline to evaluate a number of key
    issues including strand imbalance, library complexity and
    enrichment. Assessment of these characteristics are facilitated
    through diagnostic plots and summary statistics calculated over
    regions of the genome with varying levels of coverage.

    The pipeline explores and quantifies these aspects by partitioning
    the experiment reads into a collection of regions, calculating a
    series of summary statistics for each region, providing
    visualizations and calculating measures to globally asses the
    quality of a ChIP-exo experiment. We provide guidelines to
    distinguish libraries with low complexity from deeply sequenced
    experiments by the use of the QC pipeline. We demonstrate that the
    ChIP-exo QC pipeline it is also applicable to ChIP-nexus data,
    showing that those experiments present higher quality than
    ChIP-exo experiments under similar conditions.

    We compared ChIP-exo with Paired End (PE) and Single End (SE)
    ChIP-Seq and found the following characteristics: First, although
    often assumed in ChIP-exo data analysis methods, the ``peak pair''
    assumptions does not hold locally in actual ChIP-exo data. Second,
    we for the first time compared PE ChIP-Seq with ChIP-exo and found
    that at fixed sequencing depths, ChIP-exo provides higher
    sensitivity, specificity and spatial resolution than PE
    ChIP-Seq. Finally, we show that ChIP-exo and PE ChIP-Seq are
    comparable in sensitivity for closely located binding events, but
    as the average distance between binding events increases, ChIP-exo
    show higher sensitivity than PE ChIP-Seq.

    % both protocols are comparable in resolutions and sensitivity for
    % closely located binding events, but as the distance between
    % binding events increases ChIP-exo shows higher sensitivity that
    % PE ChIP-Seq. Finally, at fixed sequencing depths, ChIP-exo
    % provides higher sensitivity, specificity and spatial resolution
    % than PE ChIP-Seq.
  \end{abstract}

@@

\color{blue}
